Gradual Soundness: Lessons from Static Python

### 2022-04-04 pj-7.1/paper.tex
pj: note for later, ack Meta funding
pj: update speedup %s, follow up in microbench section

\begin{itemize}
  \item
    \emph{Evidence}.
    We present evidence that \SP{} improves performance with few code changes.
    At Instagram, the web server has become significantly more efficient
    after the application of \SP{} to high-profile modules.
    Because the server code is closed-source, we additionally present data for
    \numbenchmark{} public microbenchmarks~(\cref{s:microbenchmarks}).
    Using only shallow types, \SP{} runs slightly faster than Python
    despite the costs of enforcing soundness~(avg. \pctnum{28}).
    With fine-tuned types, \SP{} consistently outperforms the Python baseline~(avg. \pctnum{67}).

@@

Almost all ratios are less than one, showing that \SP{} achieves a speedup over
Python.
The exceptions are the simply-typed versions of \bmname{fannkuch} and
\bmname{nbody}, which is unsurprising because these benchmarks focus
on number-crunching and \SP{} optimizes arithmetic only for primitive types.
Indeed, changing a few types in \bmname{fannkuch} to primitives
gives an immediate speedup (ratio = $0.90$).

@@

%\acks{
%  \SP{} and the Cinder JIT were developed by
%  Max Bernstein,
%  John Biesnecker,
%  Jacob Bower,
%  Sinan Cepel,
%  Tiansi Hu,
%  Orvid King,
%  Vladimir Matveev,
%  William Meehan,
%  Carl Meyer,
%  Matt Page,
%  Aniket Panse,
%  Brett Simmers,
%  Andrei Talaba,
%  Dino Viehland,
%  and Shiyu Wang.
%  The Brown authors contributed to the language by modeling it independently; they
%  are not part of the \SP{} team.
%
%  Special thanks to Twitter, thanks to which the two teams
%  (specifically, Meyer and Krishnamurthi) met.
%  This work was partly supported by the US National Science Foundation.
%  Greenman acknowledges support from NSF grant 2030859 to the CRA for the \href{https://cifellows2020.org}{CIFellows} project.
%}

>>>>

\begin{itemize}
  \item
    \emph{Evidence}.
    We present evidence that \SP{} improves performance with few code changes.
    At Instagram, the web server has become significantly more efficient
    after the application of \SP{} to high-profile modules.
    Because the server code is closed-source, we additionally present data for
    \numbenchmark{} public microbenchmarks~(\cref{s:microbenchmarks}).
    Using only shallow types, \SP{} runs slightly faster than Python
    despite the costs of enforcing soundness~(avg. \pctnum{20}).
    With fine-tuned types, \SP{} consistently outperforms the Python
    baseline~(avg. \pctnum{70}).

@@

Almost all ratios are less than one, showing that \SP{} achieves a speedup over
Python (typed avg. $0.8$, refined avg. $0.3$).
The exceptions are the unrefined versions of \bmname{fannkuch} and
\bmname{nbody}, which is because these benchmarks focus
on number-crunching and \SP{} optimizes arithmetic only for primitive types.
Changing a few types in \bmname{fannkuch} to primitives
gives an immediate speedup (ratio = $0.90$).

@@

%\acks{
%  \SP{} and the Cinder JIT were developed by
%  Max Bernstein,
%  John Biesnecker,
%  Jacob Bower,
%  Sinan Cepel,
%  Tiansi Hu,
%  Orvid King,
%  Vladimir Matveev,
%  William Meehan,
%  Carl Meyer,
%  Matt Page,
%  Aniket Panse,
%  Brett Simmers,
%  Andrei Talaba,
%  Dino Viehland,
%  and Shiyu Wang.
%  The Brown authors contributed to the language by modeling it independently; they
%  are not part of the \SP{} team.
%
%  Special thanks to Twitter, thanks to which the two teams
%  (specifically, Meyer and Krishnamurthi) met.
%  This work was partly supported by the US National Science Foundation.
%  Greenman acknowledges support from NSF grant 2030859 to the CRA for the \href{https://cifellows2020.org}{CIFellows} project.
%  TODO partial funded by Meta
%}
